The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey

### 2020-05-15 RET_Hosts_CWR.tex

\replyrick{The Dark Energy Survey (DES) makes use of the Dark Energy Camera (DECam; \citealt{Flaugher2015}) to survey $5000\deg^2$ of the southern sky between 2013 and 2019. The supernova programme (DES-SN), designed primarily to perform cosmology with type Ia SNe \citep{Bernstein2012}, consisted of five 6-month seasons of approximately 7 day cadence \citep{DAndrea2018}} in ten single pointings, known as the SN fields. Of these fields, two were observed with longer exposure times and are referred to as deep fields; the remaining eight are labelled shallow. Transient detection was performed by difference imaging using a custom pipeline \citep{Kessler2015}. Transient candidates were vetted via machine-learning techniques \citep{Goldstein2015}, leading to $\sim 30,000$ viable supernova candidates over the full five seasons. A full description of the search for rapid transients in DES can be found in \citetalias{Pursiainen2018}.

@@

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typcially active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SNe-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (391 SNe and 79 AGN). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process.

\subsubsection{Final DES RET sample}
The GP lightcurves of all DES-SN candidates are classified as AGN-like or SNe-like by the CNN, and are then subject to the lightcurve quality cuts i) and ii) (Section \ref{subsubsec:trise_gr}), resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined \replyrick{for} RETs. These objects are subject to a further set of cuts \replyrick{in order to remove remaining contaminants}. We impose a cut based on a fit of the lightcurve with the \texttt{PSNID} software \citep{Sako2008}. \replyrick{\texttt{PSNID} returns a normalised goodness of fit to different SN templates, along with a probability of it being a SN Ia given that it is a SN. To remove highly-probably SNe Ia, we use thresholds of $<0.91$ and $<0.82$ respectively to the above algorithms,} which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected by MP, with the majority rejected for being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing evidence for a longer timescale decline.

@@

@@

\section{Analysis}
\label{sec:analysis} % used for referring to this section from elsewhere

@@

Another possibility is that the hosts without a redshift are mostly non star-forming, passive galaxies, for which a redshift is typically harder to obtain than for emission-line galaxies \citep{Yuan2015,Childress2017,Lidman2020}. To test this possibility, we examined the RETs that do not have a host galaxy redshift. Table \ref{tab:z_cuts} shows the numbers of RETs that failed various stages of the redshifting process, and is summarised in Fig. \ref{fig:mag_dist}. Of the 57 objects without a redshift, 47 of them have host galaxies detected in the SN Deep coadds of \citetalias{Wiseman2020}. Of more significance is that only 40 have host galaxies in the SVA1 catalogues which were used for targetting during the OzDES campaign. The other, `hostless', objects are either transients that are located remotely from a galaxy that was detected, or are hosted by a galaxy that was not detected. Non-detected hosts are either intrinsically faint and thus low in mass, situated at high redshift, or both. Neither are expected to be systematically higher in metallicity than the detected hosts. Similarly, a further 22 hosts were detected but not targetted by OzDES, due to being too faint to pass the selection criteria ($m_r < 24.5$), leaving 18 that were targetted but no redshift was found. The resulting redshift completeness of targetted objects is 71\% (83\% for objects brighter than $m_r = 24$~mag), which is in line with the average for OzDES \citep{Lidman2020}.
In Fig. \ref{fig:g-i} we show the observer-frame $r$-band magnitudes and $g-i$ colours for all RET hosts that were detected. The 47 objects with detected hosts but no redshift lie at fainter magnitudes, and appear to extend to bluer colours than those with secure redshifts. This is contrary to the hypothesis that they are high-redshift and/or passive hosts, but instead are low-mass, star-forming galaxies whose line fluxes were not strong enough to be detected.

>>>>

\replyrick{The Dark Energy Survey (DES) makes use of the Dark Energy Camera (DECam; \citealt{Flaugher2015}) to survey $5000\deg^2$ of the southern sky between 2013 and 2019. The supernova programme (DES-SN), designed primarily to detect type Ia SNe (SNe Ia) for cosmological measurements \citep{Bernstein2012}, consisted of five 6-month seasons of approximately 7 day cadence \citep{DAndrea2018}} in ten single pointings, known as the SN fields which cover a total area of $\sim27 \deg^2$. Of these fields, two were observed with longer exposure times and are referred to as deep fields; the remaining eight are labelled shallow. Transient detection was performed by difference imaging using a custom pipeline \citep{Kessler2015}. Transient candidates were vetted via machine-learning techniques \citep{Goldstein2015}, leading to $\sim 30,000$ viable supernova candidates over the full five seasons. A full description of the search for rapid transients in DES can be found in \citetalias{Pursiainen2018}.

@@

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typically active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SNe-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (391 SNe and 79 AGN). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process.

\subsubsection{Final DES RET sample}
The GP lightcurves of all DES-SN candidates are classified as AGN-like or SNe-like by the CNN, and are then subject to the lightcurve quality cuts i) and ii) (Section \ref{subsubsec:trise_gr}), resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined \replyrick{for} RETs. These objects are subject to a further set of cuts \replyrick{in order to remove remaining contaminants}. We impose a cut based on a fit of the lightcurve with the \texttt{PSNID} software \citep{Sako2008}. \replyrick{\texttt{PSNID} returns a normalised goodness of fit to different SN templates, along with a probability of it being a SN Ia given that it is a SN. To remove highly-probably SNe Ia, we use thresholds of $<0.91$ and $<0.82$ respectively to the above algorithms,} which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected, with the majority rejected for \replyrick{clearly} being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing a longer timescale decline.

@@

\begin{table}
    \centering
    \caption{\replyrick{Summary statistics for the samples compared in this work. Here we show the mean and standard deviation of the stellar mass, sSFR, and oxygen abundance, but note that the values for these parameters are not necessarily normally distributed for each sample.} }
    \begin{tabular}{lccc}
    \toprule
   Sample & $\langle\log \left(M_*\right)\rangle$ &  $\langle\log \left(\mathrm{sSFR}\right)\rangle$ &$\langle12+\log\left(\mathrm{O/H}\right) \rangle$\\
   \midrule
   {} & $\left( M_{\odot}\right)$ & $\left(\mathrm{yr}^{-1} \right)$& \\
   \midrule
   DES RETs &$9.9 \pm 0.8$ & $-9.6 \pm 0.6 $ &$ 8.4 \pm 0.2$ \\
    PS1 RETs &$9.0 \pm 0.8$ & $-9.2 \pm 0.8 $ &$ 8.5 \pm 0.2$ \\
    SNe II &$9.8 \pm 0.8$ & $-9.8 \pm 0.7 $ &$ 8.6 \pm 0.2$ \\
    SESNe &$9.5 \pm 0.9$ & $-10.2 \pm 1.0 $ &$ 8.4 \pm 0.2$ \\
    LGRBs &$9.2 \pm 0.6$ & $-8.9 \pm 0.8 $ &$ 8.4 \pm 0.2$ \\
    SLSNe &$8.5 \pm 0.9$ & $-8.7 \pm 0.9 $ &$ 8.3 \pm 0.2$ \\
\bottomrule
    \end{tabular}

    \label{tab:summary}
\end{table}

@@

\section{Analysis}
\label{sec:analysis} % used for referring to this section from elsewhere
\replyrick{A summary of the masses, sSFRs and metallicities for the hosts of DES RETs and various comparison samples is presented in Table \ref{tab:summary}. In the following sections, we compare the host galaxy properties of RETs to each sample in detail. }

@@

Another possibility is that the hosts without a redshift are mostly non star-forming, passive galaxies, for which a redshift is typically harder to obtain than for emission-line galaxies \citep{Yuan2015,Childress2017,Lidman2020}. To test this possibility, we examined the RETs that do not have a host galaxy redshift. Table \ref{tab:z_cuts} shows the numbers of RETs that failed various stages of the redshifting process, and is summarised in Fig. \ref{fig:mag_dist}. Of the 57 objects without a redshift, 47 of them have host galaxies detected in the SN Deep coadds of \citetalias{Wiseman2020}. Of more significance is that only 40 have host galaxies in the SVA1 catalogues which were used for targetting during the OzDES campaign. The other, `hostless', objects are either transients that are located remotely from a galaxy that was detected, or are hosted by a galaxy that was not detected. Non-detected hosts are either intrinsically faint and thus low in mass, situated at high redshift, or both. Neither are expected to be systematically higher in metallicity than the detected hosts. Similarly, a further 22 hosts were detected but not targetted by OzDES, due to being too faint to pass the selection criteria ($m_r < 24.5$), leaving 18 that were targetted but no redshift was found. The resulting redshift completeness of targetted objects is 71\% (83\% for objects brighter than $m_r = 24$~mag), which is in line with the average for OzDES \citep{Lidman2020}.
In Fig. \ref{fig:g-i} we show the observer-frame $r$-band magnitudes and $g-i$ colours for all RET hosts that were detected. The 47 objects with detected hosts but no redshift lie at fainter magnitudes, and appear to extend to bluer colours than those with secure redshifts. This is contrary to the hypothesis that they are high-redshift and/or passive hosts, but instead are low-mass, star-forming galaxies whose line fluxes were not strong enough to be detected. \replyrick{We thus believe that the results presented from the subset of hosts with measured redshifts are at the very least representative of the star-forming nature of the population of RET hosts}.
